### 2015-02-27 icfp15-servo/servo.tex
More more nitpicking

\subsection{Compilation strategy}
Many statically typed implementations of polymorphic languages such as Standard ML of New Jersey~\cite{SMLNJ} and
\ocaml{}~\cite{ocaml-manual-3.0} have used a compilation strategy that optimizes representations of data types when
polymorphic code is monomorphically used, but defaults to a less efficient style otherwise, in order to share
code~\cite{ocaml-repr}.
This strategy reduces code size, but leads to unpredictable performance and code, as changes to a codebase that
either add a new instantiation of a polymorphic function at a given type or, in a modular compilation setting, that
expose a polymorphic function externally, can change the performance of code that is not local to the change being
made.
Monomorphization, as in MLton~\cite{mlton-compiler}, instead instantiates each polymorphic code block at each of the types
it is applied agains, and providing predictable output code to developers, at the cost of some code duplication.
This approach, combined with whole-program compilation~\cite{weeks:whole-program-mlton}, has a negative impact on
compilation speed but has resulted in Rust code that easily matches the sequential speed of its \Cplusplus{} analog
without requiring the Servo developers to become compiler experts.

@@

\subsection{Libraries and abstractions}
Many high-level languages provide abstractions over I/O, threading, parallelism, and concurrency.
This area is one where we started out with much larger aspirations in Rust but ultimately had to pare back,
for reasons very similar to the reasons that we have avoided GC by default --- it is difficult to provide a
predictable, fast implementation that works across all platforms.
For example, thread-local storage (TLS) provides variables that are associated with a given thread.
In the presence of both lightweight (stackless) and full native threads across a variety of architectures, though,
it becomes difficult to understand what the cost of initialization and storage are for this feature.
Even lightweight threads in general, when interacting with a built-in scheduler and generic I/O system that
has to work across a variety of platforms (including Windows!) were too expensive to support in practice and
had to be removed.

Therefore, much like other modern browsers, Servo contains many of its own specialized implementations of
library functions that are tuned for the specific cases of web browsers.
For example, we have special ``small'' vectors that allow instantiation with a default inline size,
as there are uses cases where we create many thousands of vectors, nearly none of which have more than 4~elements.
In that case, removing the extra pointer indirection --- particularly if the values are of less than pointer size ---
can be a significant space savings.
We also have our own work-stealing library that has been tuned to work on the flow tree during the process of layout,
as described in \secref{sec:browsers}.
It is our hope that this code might be useful to other projects as well, though it is fairly browser-specific today.

>>>>

\subsection{Compilation strategy}
Many statically typed implementations of polymorphic languages such as Standard ML of New Jersey~\cite{SMLNJ} and
\ocaml{}~\cite{ocaml-manual-3.0} have used a compilation strategy that optimizes representations of data types when
polymorphic code is monomorphically used, but defaults to a less efficient style otherwise, in order to share
code~\cite{ocaml-repr}.
This strategy reduces code size, but leads to unpredictable performance and code, as changes to a codebase that
either add a new instantiation of a polymorphic function at a given type or, in a modular compilation setting, that
expose a polymorphic function externally, can change the performance of code that is not local to the change being
made.
Monomorphization, as in MLton~\cite{mlton-compiler}, instead instantiates each polymorphic code block at each of the types
it is applied against, providing predictable output code to developers at the cost of some code duplication.
This approach, combined with whole-program compilation~\cite{weeks:whole-program-mlton}, has a negative impact on
compilation speed but has resulted in Rust code that easily matches the sequential speed of its \Cplusplus{} analog
without requiring the Servo developers to become compiler experts.

@@

\subsection{Libraries and abstractions}
Many high-level languages provide abstractions over I/O, threading, parallelism, and concurrency.
This area is one where we started out with much larger aspirations in Rust but ultimately had to pare back,
for reasons very similar to the reasons that we have avoided GC by default --- it is difficult to provide a
predictable, fast implementation that works across all platforms.
For example, thread-local storage (TLS) provides variables that are associated with a given thread.
In the presence of both lightweight (stackless) and full native threads across a variety of architectures, though,
it becomes difficult to understand what the cost of initialization and storage are for this feature.
Even lightweight threads in general, when interacting with a built-in scheduler and generic I/O system that
has to work across a variety of platforms (including Windows!), were too expensive to support in practice and
had to be removed.

Therefore, much like other modern browsers, Servo contains many of its own specialized implementations of
library functions that are tuned for the specific cases of web browsers.
For example, we have special ``small'' vectors that allow instantiation with a default inline size,
as there are use cases where we create many thousands of vectors, nearly none of which have more than 4~elements.
In that case, removing the extra pointer indirection --- particularly if the values are of less than pointer size ---
can be a significant space savings.
We also have our own work-stealing library that has been tuned to work on the flow tree during the process of layout,
as described in \secref{sec:browsers}.
It is our hope that this code might be useful to other projects as well, though it is fairly browser-specific today.
